---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Anaelle Gackiere"
date: Sys.Date()
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
# install.packages("devtools")
# devtools::install_github("rstudio-education/dsbox") did this for myself
library(tidyverse) 
library(dsbox) 
data ("dennys")
data ("laquinta")
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

There are `r nrow(dennys)` rows and `r ncol(dennys)` columns for the Dennys dataset.Each row represents one Dennys location and the variables include city, address, state, zip, longitude, and latitude. 

```{r dennys-description}
dim(dennys)

head(dennys)
```


### Exercise 2

```{r laquinta-description}
dim(laquinta) # just checking

head(laquinta)
```

There are `r nrow(laquinta)` rows and `r ncol(laquinta)` columns for the La Quinta dataset.
Each row represents one La Quinta location and the variables include city, address, state, zip, longitude, and latitude. 

### Exercise 3
On the Dennys website, I only see US locations. On the La Quinta website, I saw that there are locations in Canada, Mexico, China, New Zealand, Georgia, Turkey, United Arab Emirates, Columbia, and Ecuador.


### Exercise 4
One way to determine whether or not there are establishments outside the United States would be by filtering the locations.

### Exercise 5 and 6
```{r locations-NonUS-DN}

# For Dennys
dennys %>%
  filter(!(state %in% states$abbreviation))

dennys %>%
  mutate(country = "United States")

```


### Exercise 7 and 8
```{r locations-NonUS-LQ}

# For La Quinta
laquinta %>%
  mutate(country = case_when(
    state %in% state.abb ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT" ~ "Colombia",
    ... # fill in the rest
  ))

laquinta <- laquinta %>%
  filter(country == "United States")


```

### Exercise 9

Which states have the most and fewest Denny’s locations? What about La Quinta? Is this surprising? Why or why not?

```{r}
# states w most and least dennys



# states w most and least laquintas



# Denny’s locations per thousand square miles.
dn %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))
```
...

### Exercise 6

Which states have the most Denny’s locations per thousand square miles? What about La Quinta?

```{r}

dn <- dn %>%
  mutate(establishment = "Denny's")
lq <- lq %>%
  mutate(establishment = "La Quinta")

```


...

 
### Exercise 7

...

### Exercise 8

...

### Exercise 9

...

### Exercise 10

...

### Exercise 11

...

### Exercise 12

...

### Exercise 13

...







